A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management

### 2022-07-03 papers/earths-future/submission.tex
Klaus handwritten edits, some questions pending

\documentclass[draft]{agujournal2019}

@@

% ACRONYMS
\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gev}{GEV}{generalized extreme value}
\newacronym{hazus}{HAZUS}{Hazard U.S.}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
\newacronym[]{usgs}{USGS}{United States Geological Survey}
\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}

@@

\draftfalse

@@

\begin{keypoints}
  \item Decisions about how to manage climate risks are often made under deep uncertainty
  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management
\end{keypoints}

@@

What are sound strategies to manage risks driven by climatic changes?
Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
Different experts can disagree about these choices, leading to different plausible projections.
Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
At one extreme, picking a single representative projection can lead to under-estimation of risk and poor decisions.
At the other, communicating results separately for each projection can overwhelm decision-makers.
To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
We develop and demonstrate a framework to address these challenges.
The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller samples and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.

@@

Many critical infrastructure services around the world are aging and inadequate, and anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet societal needs \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,ho_dams:2017}.
The need to upgrade existing infrastructure and build new facilities motivates the question: for which possible futures should infrastructure systems and components be designed?
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines.
For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily on standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, informed by probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
Additionally, local building codes \cite<based on guidance such as>[]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
Probabilistic analysis also informs the design of large-scale infrastructure.
For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
However, there are also limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.

The approahces currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.

@@

Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

@@

Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and can often be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.

@@

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose \cite{bernardo_bayesian:1994}.
Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model is reached \cite{gelman_workflow:2020}.

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \cite{quinn_exploratory:2020}.

@@

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite{schneider_scenarios:2002}.

@@

The challenge then becomes suitable estimation of the $w_j$.
Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.

@@

Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the freely available NOAA tides and currents data archive \cite{noaa_tidesandcurrents:2022}.
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
For each calendar year we first remove the annual mean, then calculate the maximum water level.
We refer this time series of annual maximum storm surges as $y'(t)$.
We display this time series of annual maxima storm surges in \cref{fig:surge-obs-return}(a).
The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.

@@

For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
We evaluate the model's fit using posterior predictive checks \cite<see>[section 2.4 and references therein]{gelman_workflow:2020}.
Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \cite<as in>[]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.

@@

We define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
The expected annual damage is thus
\begin{equation}\label{eq:ead}
  \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'},
\end{equation}
where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
Following \citeA{zarekarizi_suboptimal:2020}, we use the we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.

The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \cite<\eg,>[]{vandantzig_dike:1956}.
However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).

@@

We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.
We can also compare to the analyses of \citeA{sriver_sealevel:2018} which uses a rescaled Beta distribution with bounds of \SIrange{0.83}{8.2}{ft} and a most plausible estimate of \\SI{3.1}{ft}.
Our samples bound all of these estimates.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{scenario-map-slr-cost}
  \caption{
    Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
    Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
    The lowest-cost outcomes occur when exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
    The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
    In all cases, elevating the house reduces the variance in total lifetime cost.
    Values are sensitive to model constants; see \cref{tab:uncertainties}.
  }\label{fig:scenario-map-slr-cost}
\end{figure}

>>>>

\documentclass{agujournal2019}

@@

% ACRONYMS
\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gev}{GEV}{generalized extreme value}
\newacronym{hazus}{HAZUS}{Hazard U.S.}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{msl}{MSL}{mean relative sea level}
\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{slr}{SLR}{sea level rise}
\newacronym{ssp}{SSP}{shared socio-economic pathway}
\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
\newacronym[]{usgs}{USGS}{United States Geological Survey}
\newacronym[\glslongpluralkey={states of the world}]{sow}{SOW}{state of the world}

@@

%\draftfalse

@@

\begin{keypoints}
  \item Decisions about how to manage climate risks are often made under deep uncertainty.
  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
\end{keypoints}

@@

What are sound strategies to manage risks driven by climatic changes?
Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
Different experts can disagree about these choices, leading to different plausible projections.
Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
At one extreme, picking a single representative projection can lead to under-estimation of risk and poor decisions.
At the other extreme, communicating results separately for each projection can overwhelm decision-makers.
To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
We develop and demonstrate a framework to address these challenges.
The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller samples and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.

@@

Many critical infrastructure services around the world are aging \cite{ho_dams:2017}.
Further, anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet their design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines \cite{bruneau_multihazard:2017}.
For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings \cite{asce_7-10:2013}.
At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.

@@

Current practice in engineering, infrastructure design, and local governance relies heavily on standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
These are often, though not always, informed by probabilistic analysis of relevant data.
For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
Additionally, local building codes (based on guidance such as \citeA{asce_24-05:2006} or \citeA{FEMA_p-55:2011}) may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
Probabilistic analysis also informs the design of large-scale infrastructure.in particu
For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.

There are many advantages to standards-based design.
In particular, these heuristics are scalable and explainable, they reduce the complexity of design analysis, and they are fair in at least a procedural sense.
However, reliance on these heuristics also faces limitations.
In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and can be particularly sensitive to representation of tail probabilities \cite{merz_heavytails:2022,wong_floodrisk:2018,garner_slrise:2018}.

The approaches currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.\footnote{Klaus comment TBD}
While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,Montanari:2014hl,Serinaldi:2015bq}.
Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.

@@

Projecting nonstationary hazard is difficult because many future hazards depend on human decisions (\eg, future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.

@@

Although these \gls{dmdu} methods have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and can often be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.

@@

We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.

@@

A first analytical step is to use the model in an ``exploratory'' mode.
Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020}.

@@

Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite<see>{schneider_scenarios:2002}.

@@

The challenge then becomes to suitably estimate the $w_j$.
Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.

@@

We use data on storm surge at Sewells Point, VA (gauge 8638610) from the \gls{noaa} tides and currents data archive \cite{noaa_tidesandcurrents:2022}.
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
For each calendar year we first remove the annual mean, then calculate the maximum water level.
We refer this time series of annual maximum storm surges as $y'(t)$.
We display this time series of annual maxima storm surges in \cref{fig:surge-obs-return}(a).
The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.

@@

For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{betancourt_hamiltonian:2018,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
We evaluate the model's fit using posterior predictive checks \cite<see>[section 2.4 and references therein]{gelman_workflow:2020}.
Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \cite<as in>[]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.

@@

We define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
The expected annual damage is thus
\begin{equation}\label{eq:ead}
  \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'},
\end{equation}
where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.

The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \cite<\eg,>[]{vandantzig_dike:1956}.
However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution.
Instead, we estimate this convolution through a Monte Carlo method (see \cref{sec:alg-ead} for details).
Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).

@@

We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.
We can also compare to the analyses of \citeA{sriver_sealevel:2018} which uses a rescaled Beta distribution with bounds of \SIrange{0.83}{8.2}{ft} and a most plausible estimate of \SI{3.1}{ft}.
Our samples bound all of these estimates.

@@

\begin{figure}
  \includegraphics[width=\textwidth]{scenario-map-slr-cost}
  \caption{
    Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \acrfull{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
    Colors indicate the number of \acrfullpl{sow} of falling within each box.
    The lowest-cost outcomes occur when exposure is low ($h_0$ is large and \acrfull{slr} is minimal) and the house is not elevated (no up-front cost).
    The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
    In all cases, elevating the house reduces the variance in total lifetime cost.
    Values are sensitive to model constants; see \cref{tab:uncertainties}.
  }\label{fig:scenario-map-slr-cost}
\end{figure}

One application of exploratory analysis is to reveal the range and variation in outcomes, conditional on taking a particular decision.
\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
The outcomes with lowest total lifetime costs arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
